---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
library(ggpubr)
```


```{r}
# Us this if you need to load new data, we are using the output of this saved as an RDS below
# Biomark chip data
# W1_1_Raw_Data <- read_excel("~/Desktop/biomark-private/data/W1.1 Raw Data.xlsx", skip = 11)
# colnames(W1_1_Raw_Data)
# W1_2_Raw_Data <- read_excel("~/Desktop/biomark-private/data/W1.2 Raw Data.xlsx", skip = 11)
# W2_1_Raw_Data <- read_excel("~/Desktop/biomark-private/data/W2.1 Raw Data.xlsx", skip = 11)
# W3_1_Raw_Data <- read_excel("~/Desktop/biomark-private/data/W3.1 Raw Data.xlsx", skip = 11)
# Raw_Data <- rbind(W1_1_Raw_Data, W1_2_Raw_Data, W2_1_Raw_Data, W3_1_Raw_Data) %>%
#   select("Sample_Name" = Name...2, "Sample_Type" = Type...3, "Assay" = Name...5, "Assay_Type" = Type...6, everything())
```

```{r}
# Worle meta data
# Worle_biomark_meta <- read.csv("~/Desktop/biomark-private/meta/Worle_biomark_meta.csv")
# Worle_biomark_meta <- Worle_biomark_meta %>%
#   select("Sample_Name" = Name, everything())
```

```{r}
# # compare names 
# bmk <- as.character(unique(Worle_biomark_meta$Sample_Name))
# smp <- unique(Raw_Data$Sample_Name)
# intersect(bmk, smp)
# # in bmk but not smp
# setdiff(bmk, smp) 
# # in smp but not bmk
# setdiff(smp, bmk) 
```
```{r}
# # fix names
# # remove _STD and rename intl to intI
# Data <- Raw_Data %>%
#   mutate_all(~gsub("_STD", "", .)) %>% 
#   mutate_all(~gsub("intl1F165_clinical", "intI1F165_clinical", .)) 
# bmk <- as.character(unique(Worle_biomark_meta$Sample_Name))
# smp <- unique(Data$Sample_Name)
# 
# # in bmk but not smp
# setdiff(bmk, smp) 
# # in smp but not bmk
# setdiff(smp, bmk) 
```
```{r}
# merge
# df <- left_join(Data, Worle_biomark_meta)
# class(df)
# rm(Raw_Data, W1_1_Raw_Data, W1_2_Raw_Data, W2_1_Raw_Data, W3_1_Raw_Data, Worle_biomark_meta, Data, bmk, smp)
# 
# # reorder sample_day column
# df$sample_day <- factor(df$sample_day, levels=c("TB", "T00", "T02", "T14"))
# saveRDS(df, file = "../data/metawithbiomark.RDS")
df <- readRDS("../data/metawithbiomark.RDS") 
```

```{r}
# subset biomark chip to interesting things
soil <- df %>%
  filter(sample_type %in% "soil")
water <- df %>%
  filter(sample_type %in% "water") 
manure <- df %>%
  filter(sample_type %in% "manure")
control <- df %>%
  filter(sample_type %in% "control")
standard <- df %>%
  filter(sample_type %in% "standard")
```
# Look at manure
What ARGs did we detect in manure? 
What ARGs did we detect in non-manured soil plots? Should we ignore those for this analysis? 
```{r}
m <- manure %>%
  filter(Value < 28, Call == "Pass") %>%
  group_by(sample_type, Assay) %>%
  dplyr::mutate(mean = mean(as.numeric(Value)))

manureplot <- m %>%
  ggplot(aes(x = reorder(Assay, -mean), y = mean, fill = Assay)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_viridis_d() +
    coord_flip() +
    labs(y = "mean Ct value", x = "Assay", title = "Cycle times for assays in poultry manure \n sorted by most abundant assay on top")
manureplot

order <- m %>%
  group_by(sample_type, Assay) %>%
  arrange(-mean) %>%
  distinct(Assay) %>%
  pull(Assay)

plot <- manure %>%
  filter(Value < 28, Call == "Pass") %>%
  group_by(Sample_Name, Assay) %>%
  dplyr::mutate(mean = mean(as.numeric(Value))) %>%
  ggboxplot(x = "Assay", y = "mean", order = order, fill = "Assay") +
    scale_fill_viridis_d() +
    coord_flip() +
    theme(legend.position = "none") +
    labs(y = "mean Ct value", x = "Gene", title = "Cycle times for genes of interest in poultry manure \n sorted by most abundant gene on top")
plot
ggsave(filename = "../plots/pultrymanureCT.png", plot = last_plot(), device = "png")

```
Non-manured soils, baseline
```{r}
m <- soil %>%
  filter(Value < 28, Call == "Pass", sample_day == "TB") %>%
  group_by(sample_type, Assay) %>%
  dplyr::mutate(mean = mean(as.numeric(Value)))

soilplot <- m %>%
  ggplot(aes(x = reorder(Assay, -mean), y = mean, fill = Assay)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_viridis_d() +
  coord_flip() +
  labs(y = "mean Ct value", x = "Assay", title = "Cycle times for assays in non-manured soils \n sorted by most abundant assay on top")
soilplot
```
Non-manured soils, no manure treatment
```{r}
m <- soil %>%
  filter(Value < 28, Call == "Pass", treatment == "WCS") %>%
  group_by(sample_type, Assay) %>%
  dplyr::mutate(mean = mean(as.numeric(Value)))

soilplot2 <- m %>%
  ggplot(aes(x = reorder(Assay, -mean), y = mean, fill = Assay)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_viridis_d() +
    coord_flip() +
    labs(y = "mean Ct value", x = "Assay", title = "Cycle times for assays in non-manured soils \n sorted by most abundant assay on top")
soilplot2
```
```{r}
list <- unique(m$Assay)
list

for (i in 1:length(list)) {
  count <- plot_standard_curve_counts(standard, "i")
  print(list[i])
}
count.data <- count$data

```

```{r}
print("These assays are in manure but not non-manured soil")
manargs <- setdiff(manureplot$data$Assay, soilplot$data$Assay)
manargs
```
Based on above, we should choose some interesting genes that are in manure but not soil.

16s, tetH, tetM, tetT, tetW, tetX, erm(C)

What about soil associated ARGs? Those ARGs that are not from manure? 

```{r}
print("These assays are in soil but not manure")
soilargs <- setdiff(soilplot$data$Assay, manureplot$data$Assay)
soilargs
```

We need to calculate count using this function:
`count = SUM(B2)*(1/(10^9))*(1/660)*((6.023*10^23)/1)*(1/C2)*(D2)`

B2 = concentration (decimal)
C2 = amplicon length (varies, see cybox biomark for each assay)
D2 = sample volume in ul (always .0067 for biomark)

For each of the genes of interest

```{r}
# File of amplicon lengths (phil's biomark folder on Box)
PC_Standards <- read_excel("~/Downloads/PC8_Standards.xlsx")
colnames(PC_Standards)
# rename some columns
PC_Standards <- PC_Standards %>%
  select("gene" = `Sequence Name`, "amp_len" = `Total size`)

head(PC_Standards)
# 16s Assay names need fixed from Phil's to match biomark
PC_Standards_no_NA <- PC_Standards %>% 
  separate(gene, c("Assay"), sep = -2, remove = F) %>% select(gene, Assay, amp_len) %>%
  mutate_all(~gsub("16S_Eub_338", "16S_Eub_338_518", .)) %>%
  drop_na()

head(PC_Standards_no_NA)
tail(PC_Standards_no_NA)
rm(PC_Standards)
saveRDS(PC_Standards_no_NA, "../data/ampliconlengths.RDS")
PC_Standards_no_NA <- readRDS("../data/ampliconlengths.RDS")
```
```{r}
options(scipen = 999)

# plot standard curve from biomark standards using expected copy numbers based on amplicon length for each Assay
plot_standard_curve_counts <- function(standards, gene){
  
  std <- standards %>%
    filter(Value < 990, grepl(gene, Assay), grepl(gene, Sample_Name)) %>%
    left_join(PC_Standards_no_NA) %>%
    select(Assay, rConc, Value, amp_len) %>%
    mutate(counts = as.numeric(rConc) * (1/(10^9)) * (1/660) * ((6.023*10^23)/1) * (1/as.numeric(amp_len)) * (.0067)) %>%
    ggplot(aes(x = log10(counts + 1), y = as.numeric(Value))) + 
      geom_point() + geom_smooth(method = 'lm') + 
      labs(x = "Count (log10)", y = "Ct") +
      ggtitle(gene)
 
  return(std)
}
```
Plot standard curves for genes of interest, we will use the plot data later on in the analysis
```{r}
sixS <- plot_standard_curve_counts(standard, "16S")
tetH <- plot_standard_curve_counts(standard, "tetH")
tetM <- plot_standard_curve_counts(standard, "tetM")
tetT <- plot_standard_curve_counts(standard, "tetT")
tetW <- plot_standard_curve_counts(standard, "tetW")
tetX <- plot_standard_curve_counts(standard, "tetX")
tetbP <- plot_standard_curve_counts(standard, "tetbP")
strB <- plot_standard_curve_counts(standard, "strB")
#ermC <- plot_standard_curve_counts(standard, "erm(C)")
sixS 
tetH
tetM
tetT
tetW 
tetX 
strB
#ermC 
# erm(C) does not look good, leave it out. Also need to change filter function as () don't work with grepl. 
# test <- standard %>%
#   filter(Assay %in% "erm(C)", Sample_Name %in% "erm(C)") %>%
#   left_join(PC_Standards_no_NA) %>%
#   select(Assay, rConc, Value, amp_len) %>%
#   mutate(counts = as.numeric(rConc) * (1/(10^9)) * (1/660) * ((6.023*10^23)/1) * (1/as.numeric(amp_len)) * (.0067)) %>%
#   ggplot(aes(x = log10(counts + 1), y = as.numeric(Value))) + 
#     geom_point() + geom_smooth(method = 'lm') + 
#     labs(x = "Count (log10)", y = "Ct") +
#     ggtitle("erm(C)")
# test
```                                                   
```{r}                                                
# Function to make count data frame for a gene         
stdgenecnt <- function(std){                          
                                                      
  Counts <- std %>%                                   
    group_by(rConc) %>%
    summarise_at(vars(counts), funs(mean(., na.rm = TRUE))) %>%
    mutate(Assay = unique(std$Assay))
  
  return(Counts)
}
Counts16s <- stdgenecnt(sixS$data)
```

Check efficiency and fit of each standard
```{r}
dt <- sixS$data
lm.r <- lm(data = dt, formula = Value ~ log10(counts+1))
summary(lm.r)$r.squared

coef <- coef(lm.r)[2]
eff <- 100*((10^(-1/coef)) - 1)
(eff)
```
16s looks good

```{r}
#Function to calculate mean Ct value if you wish to use it
meanCT <- function(sample_data, gene){
  soil16s <- sample_data %>%
  filter(grepl(gene, Assay), Value < 28) %>%
  select(ID, Sample_Name, Assay, Value, treatment, sample_day, sample_number, soil_type, plot) %>%
  group_by(Sample_Name) %>%
  dplyr::mutate(mean = mean(Value)) #%>% # if you load `plyr` after dplyr it will not work correctly
  #select(-Value) #%>%
  #unique()
  return(soil16s)
}
```


# No ARGs in manure amplified in soil from list above. 
# Plot ARG Ct in soils 
```{r}
# search for tet genes less than 28, only return tetG_F, which didn't amplify in manure and tetL which also amplified in soil. strB not in soil either. Basically nothing detected in manure as ARG is detected in soil. Yet 16s is. 
soil16s <- soil %>%
  filter(!grepl("16S", Assay), Value < 28, Call == "Pass") %>%
  select(Sample_Name, Assay, Value, treatment, sample_day, sample_number, soil_type) %>%
  group_by(Sample_Name) %>%
  dplyr::mutate(mean = mean(Value)) #%>% # if you load `plyr` after dplyr it will not work correctly
  #select(-Value) #%>%

manargs
a <- soil16s %>%
  mutate(colssss = ifelse(Assay %in% manargs, "blue", "red")) %>%
  filter(colssss == "blue")

a %>%
  ggplot(aes(x = reorder(Assay, -mean), y = Value, fill = sample_number)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  scale_fill_viridis_d() +
  labs(y = "mean Ct value", x = "Assay", title = "Cycle times for assays in soils") +
  facet_grid(treatment ~ sample_day) +
  coord_flip()

```
The story here is that we basically are not able to detect many manure associated ARGs (those only in manure but not in un-manured soils), we must move to looking into the water. 

# Soil
```{r}
class(soil$Value)
soil$Value <- as.numeric(soil$Value)

msoil16S <- meanCT(soil, "16S")
```
plot 16s counts for depth 1
```{r}
dt <- sixS$data
dt2 <- msoil16S

inverse.lm <- lm(data = dt, formula = log10(counts+1) ~ as.numeric(Value))

val4 <- dt2$Value

dt2$sixS <- 10 ^ predict(inverse.lm, data.frame(Value = val4), interval = "predict")[,1]

d1 <- dt2 %>%
  filter(!sample_day %in% "T14") # drop random T14?

glimpse(d1)
my_comparisons <- list( c("WCS", "WCSM"), c("WCS", "WCM"), c("WCM", "WCSM"))

ggboxplot(d1, x = "treatment", y = "sixS", fill = "treatment") +
  facet_grid(sample_number ~ sample_day) +
  stat_compare_means(comparisons = my_comparisons) +
  scale_y_continuous(trans = 'log10') +
  labs(x = "Depth 1 soil samples", y = "Log 10 16s gene copies") + 
  scale_fill_viridis_d() +
  theme(legend.position = "none") 

ggsave(filename = "../plots/16scountsd1worle.png", plot = last_plot(), device = "png", width = 8, height = 8)
```


# Water

```{r}
m <- water %>%
  filter(Value < 28, Call == "Pass") %>%
  group_by(treatment, Assay, sample_number) %>%
  dplyr::mutate(mean = mean(as.numeric(Value))) %>%
  mutate(colssss = ifelse(Assay %in% manargs, "blue", "red")) %>%
  filter(colssss == "blue")

manureplot <- m %>%
  ggplot(aes(x = Assay, y = mean, fill = Assay)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_viridis_d() +
    coord_flip() +
    labs(y = "mean Ct value", x = "Assay", title = "Cycle times for assays in water") +
    facet_grid(treatment ~ sample_number)
manureplot

unique(manureplot$data$Assay)

order <- water %>%
  filter(Value < 28, Call == "Pass") %>%
  group_by(Assay) %>%
  dplyr::mutate(mean = mean(as.numeric(Value))) %>%
  arrange(-mean) %>%
  pull(Assay) %>%
  unique()

plot <- water %>%
  filter(Value < 28, Call == "Pass", Assay %in% manargs, !treatment == "WCS" ) %>%
  group_by(treatment, Assay) %>%
  dplyr::mutate(mean = mean(as.numeric(Value))) %>%
  ggboxplot(x = "Assay", y = "Value", fill = "treatment", order = order) +
    scale_fill_viridis_d() +
    coord_flip() +
    labs(y = "mean Ct value", x = "Gene", title = "Cycle times for genes from poultry manure in runoff water ")
plot
ggsave(filename = "../plots/pultrymanurenwaterCT.png", plot = last_plot(), device = "png")
```
tet genes confer various forms of resistance to the antibiotic tetracycline. 
tnp genes are involved in DNA integration and DNA recombination...potentially involved in HGT
lnuA Lincosamide nucleotidyltransferase
erm(C) Erythromycin resistance protein


After determining that we had limited detection in soil for ARGs, let's move onto the water and tet genes
```{r}
water$Value <- as.numeric(water$Value)
mwater16S <- meanCT(water, "16S")
mwatertetH <- meanCT(water, "tetH")
mwatertetM <- meanCT(water, "tetM")
mwatertetT <- meanCT(water, "tetT")
mwatertetW <- meanCT(water, "tetW")
mwatertetX <- meanCT(water, "tetX")
mwatertetbP <- meanCT(water, "tetbP")
```

Water 16s
```{r}
water$Value <- as.numeric(water$Value)

inverse.lm <- lm(data = sixS$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwater16S$Value

mwater16S$sixS <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwater16S 

ggboxplot(mwater16S, x = "sample_number", y = "sixS", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 gene 16s copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 

ggsave("../plots/16swater.png", plot = last_plot(), device = "png")
```
tetT was detected in the most water samples, 33, compared to 188 for 16s
```{r}
tetT
inverse.lm <- lm(data = tetT$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetT$Value

mwatertetT$tetT <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetT 

ggboxplot(mwatertetT, x = "sample_number", y = "tetT", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 tetT gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 

ggsave("../plots/tetTwater.png", plot = last_plot(), device = "png")
```
```{r}
tetX
inverse.lm <- lm(data = tetX$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetX$Value

mwatertetX$tetX <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetX 

ggboxplot(mwatertetX, x = "sample_number", y = "tetX", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 tetX gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 

ggsave("../plots/tetXwater.png", plot = last_plot(), device = "png")
```
```{r}
tetW
inverse.lm <- lm(data = tetW$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetW$Value

mwatertetW$tetW <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetW 

ggboxplot(mwatertetW, x = "sample_number", y = "tetW", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 tetW gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 

ggsave("../plots/tetWwater.png", plot = last_plot(), device = "png")
```
tetH
```{r}
tetH
inverse.lm <- lm(data = tetH$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetH$Value

mwatertetH$tetH <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetH 

ggboxplot(mwatertetH, x = "sample_number", y = "tetH", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 tetH gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 

ggsave("../plots/tetHwater.png", plot = last_plot(), device = "png")
```

tetM
```{r}
tetM
inverse.lm <- lm(data = tetM$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetM$Value

mwatertetM$tetM <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetM 

ggboxplot(mwatertetM, x = "sample_number", y = "tetM", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 tetM gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 
  

ggsave("../plots/tetMwater.png", plot = last_plot(), device = "png")

# divide by 16s, mean of counts by plot and time

x <- mwatertetM %>%
  dplyr::mutate(meantet = mean(as.numeric(tetM))) %>%
  select(Sample_Name, sample_number, treatment, meantet) %>%
  unique()

y <- mwater16S %>%
  dplyr::mutate(mean16s = mean(as.numeric(sixS))) %>%
  select(Sample_Name, sample_number, treatment, mean16s) %>%
  unique()

test <- left_join(x,y, by = "Sample_Name") %>%
  mutate(ncounttetM = meantet / mean16s)

ggboxplot(test, x = "sample_number.x", y = "ncounttetM", fill = "treatment.x", add = "dotplot") +
  labs(x = "Water Sample", y = "16s normalized tetM gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 
```

tetbP
```{r}
tetbP
inverse.lm <- lm(data = tetbP$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetbP$Value

mwatertetbP$tetbP <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetbP 

ggboxplot(mwatertetbP, x = "sample_number", y = "tetbP", fill = "treatment", add = "dotplot") +
  labs(x = "Water Sample", y = "Log 10 tetbP gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 

ggsave("../plots/tetbPwater.png", plot = last_plot(), device = "png")
```
# ARGs in soil but not manure, are they in water? 
```{r}
soilargs
aac3_Via <- plot_standard_curve_counts(standard, "aac3-Via")
tetG_F <- plot_standard_curve_counts(standard, "tetG_F")
cadC <- plot_standard_curve_counts(standard, "cadC")
aac3_Via
tetG_F
cadC
```
# Only tetG_F generated standard curve
* May wish to investigate why the others did not
```{r}
mwatertetG_F <- meanCT(water, "tetG_F")
tetG_F
inverse.lm <- lm(data = tetG_F$data, formula = log10(counts+1) ~ as.numeric(Value))

val2 <- mwatertetG_F$Value

mwatertetG_F$tetG_F <- 10 ^ predict(inverse.lm ,data.frame(Value = val2), interval = "predict")[,1]

mwatertetG_F 
my_comparisons <- list( c("WCS, WCSM"), c("WCS, WCM"), c("WCM, WCSM"))
ggboxplot(mwatertetG_F, y = "tetG_F", fill = "treatment", add = "jitter", shape = "treatment") +
  facet_grid(~sample_number) +
  stat_compare_means(comparisons = my_comparisons)
  labs(x = "Water Sample", y = "Log 10 tetG_F  gene copies") +
  scale_y_continuous(trans = 'log10') +
  scale_fill_viridis_d() 


ggsave("../plots/tettetG_Fwater.png", plot = last_plot(), device = "png")
```

